---
title: "P8105 Homework 2 Solutions"
author: Meghan Bellerose,
date: September 24, 2020
output: github_document
---

This homework uses r version 4.0.2 and r studio version 1.3.1073.

For this homework, I need the tidyverse and readxl libraries. 

```{r, message = FALSE}
library(tidyverse)
library(readxl)
```

# Problem 1

### Read and clean Mr. Trash Wheel data

First, I will read the Mr. Trash Wheel Excel data sheet, create snake format variable names, and omit non-data entries, including rows and columns with notes and figures, as well as entries that are unrelated to the dumpster data. 

I'll also round the number of sports balls to the nearest integer and convert the results to an integer variable.

```{r}
trashwheel_df =
  read_xlsx("./data/trash_wheel.xlsx",
  sheet = "Mr. Trash Wheel",
  range = cell_cols("A:N")) %>%
  janitor:: clean_names() %>%
  drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  ) 
```

### Read and clean precipitation data 

Next, I'll read and clean the precipitation data for 2017 and 2018, omitting rows without data and adding a variable for year. 

```{r}
precip_2018 = 
  read_excel("./data/trash_wheel.xlsx",
      sheet = "2018 Precipitation",
      skip = 1
  ) %>%
  janitor:: clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018) %>%
  relocate(year)

precip_2017 = 
  read_excel("./data/trash_wheel.xlsx",
      sheet = "2017 Precipitation",
      skip = 1
  ) %>%
  janitor:: clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017) %>%
  relocate(year)
```

Now I will combine the precipitation datasets and convert month to a character variable using month.name. 

```{r}
precip_df = 
  bind_rows(precip_2017, precip_2018) %>%
  mutate(month = month.name[month])
```

The square bracket vector is not ideal, so we can create a month data frame instead.

```{r}
options(tibble.print_min = 3)

month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

precip_df = 
  bind_rows(precip_2017, precip_2018)

left_join(precip_df, month_df, by = "month")
```
This dataset contains information from the Mr. Trashwheel collector in Baltimore, MD. As trash enters the inner harbor, the trashwheel collects the trasha and stores it in a dumpster. The dataset contains information on year, month, weight of the trash collected, and the number of some specific types of trash, such as plastic bottles. There are `r nrow(trashwheel_df)` rows in our final dataset. Additional data sheets include monthly precipitation data.  

This code can be used with the 2020 data, because the sheet names are consistent!

# Problem 2

### Read and clean the NYC Transit dataset

I will read and clean the NYC transit dataset, keeping important variables.

```{r}
options(tibble.print_min = 3)

subway_df =
  read_csv("./data/subway.csv", 
    col_types = "cccnnccccccccccccccccclclcccnncc") %>%
  janitor:: clean_names() %>%
  select(
      line, station_name, station_location, route1, route2, route3, route4, route5, route6, route7, route8, route9, route10, route11, vending, entry, entrance_type, ada)

subway_tidy_data = 
  pivot_longer(
    subway_df,
    route1:route11,
    names_to = "route_order",
    names_prefix = "route",
    values_to = "route") %>%
  drop_na(route)

subway_tidy_data


```

I then converted the entry variable from a character to a logical variable.

```{r, message = FALSE, eval = FALSE}
as.logical(pull(subway_tidy_data, entry))
```

This dataset contains information about the NYC transit systems' subway stations. I cleaned the dataset so that variable names are in snake format and only retained information on the station names and locations, the routes that are served by each station during weekday service, whether the location allows entry, the entrance type, whether vending machines are available, and whether it is ADA accessible. I tidied the data so that routes 1 through 11 were listed in a single column rather than 11 columns and rows with NA are removed. I then converted entry from a character to a logical vector. There are `r nrow(subway_df)` rows and `r ncol(subway_df)` columns in the final dataset. I consider the data to be tidy.

### NYC Transit system details

Next, I will find the number of distinct stations, the number of ADA compliant stations, and the proportion of stations without vending machines that allow entrance.

```{r}
distinct(subway_tidy_data, station_name)

filter (subway_tidy_data, ada == TRUE) %>%
  distinct(station_name)

filter (subway_tidy_data, vending == FALSE)



```
There are 356 distinct stations, and 73 are ADA compliant. 

I will now make route number and name distinct variables. 

There are X distinct stations that serve the A train. Of these, X are ADA compliant.

# Problem 3

